Low-Cost Data Deduplication for Virtual Machine Backup in Cloud Storage

### 2013-03-08 hotstorage13/abs.tex
5 papges

\begin{abstract}
%Virtualization has became the engine behind many cloud computing platforms.
In a virtualized cloud cluster, frequent  snapshot backup of virtual disks improves
hosting  reliability; howver, it  takes significant 
memory and CPU resource to perform data duplication in order to remove
excessive redudancy among image snapsots.
This paper presents a low-cost clustered deduplication soluton scalable for handling a large number 
of virtual machines.  The key idea is to seperate duplicate detection from 
the actual storage data backup instead of using inline deduplication.
Global inddex and  detection requests are partitioned 
among machines using  fingerprint values for parallel detection. 
Each machine conducts duplicate detection partiton by partition indepdently
with a minimal memory usage.
%in parallel among all machines and  each machine full duplication detection is 
Another optimization is to allocate and control
buffer space for exchanging  detection messages and duplicate summary among machines.
%The memory requirement and disk usage for the proposed solution is very small while the overall thoughput
%and backup process timing is not compromised. 
Our experiments show that  the proposed scheme 
uses a very small amount of system resources while accomplishing a satisfactory backup throughput
in a large cloud setting.
%which only reqiires a very small amount of memory and CPU resource.  
%Experimental results  show the proposed scheme  can achieve high deduplication ratio while using
%a  small  amount of cloud resources.

>>>>

\begin{abstract}
%Virtualization has became the engine behind many cloud computing platforms.
In a virtualized cloud cluster, frequent  snapshot backup of virtual disks improves
hosting  reliability; howver, it  takes significant 
memory and CPU resource to perform data duplication in order to remove
excessive redudancy among snapsots.
This paper presents a low-cost clustered deduplication soluton scalable for handling a large number 
of virtual machines.  The key idea is to seperate duplicate detection from 
the actual storage data backup instead of using inline deduplication, and partition
global index and  detection requests among machines using  fingerprint values.
Then each machine conducts duplicate detection partiton by partition indepdently with a minimal memory usage.
%in parallel among all machines and  each machine full duplication detection is 
Another optimization is to allocate and control
buffer space for exchanging  detection messages and duplicate summary among machines.
%The memory requirement and disk usage for the proposed solution is very small while the overall thoughput
%and backup process timing is not compromised. 
Our experiments show that  the proposed scheme 
uses a very small amount of system resources while accomplishing a satisfactory backup throughput
in a large cloud setting.
%which only reqiires a very small amount of memory and CPU resource.  
%Experimental results  show the proposed scheme  can achieve high deduplication ratio while using
%a  small  amount of cloud resources.

### 2013-03-08 hotstorage13/arch.tex
5 papges

\section{ Key Ideas and Design Architecture}
\label{sect:arch}

@@

\item {\bf Seperation of duplicate detection and data backup.}
%Request accumulation and partition-based deduplication.}
The second level duplicate detection requires a global comparison of 
content fingerprints of stored chunks with the new chunks scanned during 
the backup process. Bloom filters and caching allows some of content 
fingerprints to be stored on the cheap disks~\cite{bottleneck08}. Such an optimization is good 
for inline deduplication where chunks scanned need to be determined to be duplicate or not instantly without 
waiting while  it memory consumption is still significant in competing with stand virtual 
machine activities.  Our idea is to perform duplicate detection first before actual data backup.
That requres a pre-scanning of  data  blocks of VM images and after that, real backup  for those nonduplicates
is conducted. That does incur an extra  round of I/O reading for nonduplicate blocks while avoiding inline 
deduplication and leading to a much smaller resource requirement. 
During duplicate detection phase, detection requests can be accmnulated on the disk 
for a period time  in a  partitioned manner.  Aggregated duplicate requests can be processed partition by partition. 
We also accumulate the delete requests and perform them partition by partitions. 
Since each partition corresponds to a small portion of global content index, the cost to 
process requests within a partition is significant smaller than process all the requests globally.

\item {\bf Buffered data redistrubution in parallel processing of  detection  requests}.  
We let {\em global index} be the meta data containing the fingerprint values of all unique blocks
in the entire cluster and  the reference pointers which lead to the actual location of raw data.
A dupliate detection request for a content block needs to be compared with the global index to determine if this
block is nonduplicate. If it is a duplicate, return the correspinding reference pointer.
We conduct parallel processing of detection requests in all machines in a cloud cluster.
A logical way to distribute detection requests is to parititon based on the content  hashing of data blocks
and the global index can also be partitioned accordingly in a  disjointed manner.
Initial data follows the VM distribution among machines and the detected results 
represented as  duplicate summary need to be collected following VM distribution. 
Therefore, there are two all-to-all data remapping operations involved.
One is to map detection requests from VM-based requests to the fignerprint based distribution.  
One is to map dupliate summary from fingerprint-based requests to the VM based distribution.  
The redistribued data needs to be accumulated on the disk to minimize the use of memory.
To minimize the cost of disk seek, outgoing or incoming data exchange messages need to be buffered to bundle small messages.
Given there are $p\times q$ partitions where $p$ is the number of machines and $q$ is the number of fingerprint-based partitions
at each machine, the number of buffers is big and space per each buffer  is small given the memory constraint.
Then there will be a large number of storage  IO operations involved, suffering the huge seek overhead.
We have designed an effient data exchange and disk data buffering  scheme to address this.

@@

{\bf Snapshot deleteion.} Each virtual machine will keep at most $x$ snapshots of VM images and thus
expired snapshots will be deleted unless its owern explicitly asks for its retention. 
A block or the entire segment can be deleted if its reference counter is zero.
As we discussed in Section~\ref{sect:data}, the reference counter is kept in the global index, partitioned among machines. 
We deal snapshot deletion in three steps similar to the VM snapshot backup.
Step 1 is to read  the meta data of a snapshot such as segment and block
recipes and  accumate  deletion requests in different machines  in the fingerprint based distribution.
The second step  is to process the deletion requests within each partition and detect those records with zero 
reference counters. The third step is that every VM gathers the confirmed deletion requests for its cotent and sends them
to the backup storage.
The archival data repository may log these deletion requests,  and will periodically perform a compaction operation when 
its deletion log is too big. 
Due to the paper constraint, we will not discuss the deletion operations in details.

The above method requires the counter information to be consistently maintained in the global index.
Another option to avoid maitaining its consitency is to perioidcally read all receipts of active 
snapsots and compute the counter of all blocks while still following the fingerprint-based partitioning. 
%Finally the deleteion summary. it will adjust the corresponding reference counts. At the end of bucket worker's operation, it outputs the unused blocks information to each VM data repository's deletion log.

>>>>

\section{ Key Ideas and Architecture}
\label{sect:arch}

@@

\item {\bf Seperation of duplicate detection and data backup.}
%Request accumulation and partition-based deduplication.}
The second level duplicate detection requires a global comparison of 
content fingerprints of stored chunks with the new chunks scanned during 
the backup process. Bloom filters and caching allows some of content 
fingerprints to be stored on the cheap disks~\cite{bottleneck08}. Such an optimization is good 
for inline deduplication where chunks scanned need to be determined to be duplicate or not instantly without 
waiting while  it memory consumption is still significant in competing with stand virtual 
machine activities.  Our idea is to perform duplicate detection first before actual data backup.
That requres a pre-scanning of  data  blocks of VM images and after that, real backup  for those nonduplicates
is conducted. That does incur an extra  round of I/O reading for nonduplicate blocks while avoiding inline 
deduplication and leading to a much smaller resource requirement. 
During duplicate detection phase, detection requests can be accmnulated on the disk 
for a period time  in a  partitioned manner.  Aggregated duplicate requests can be processed partition by partition. 
%We also accumulate the delete requests and perform them partition by partitions. 
Since each partition corresponds to a small portion of global content index, the cost to 
process requests within a partition is significant smaller than process all the requests globally.

\item {\bf Buffered data redistrubution in parallel duplicate detection}.  
We let {\em global index} be the meta data containing the fingerprint values of all unique blocks
in the entire cluster and  the reference pointers which lead to the actual location of raw data.
A dupliate detection request for a content block needs to be compared with the global index to determine if this
block is nonduplicate. If it is a duplicate, return the correspinding reference pointer.
We conduct parallel processing of detection requests in all machines in a cloud cluster.
A logical way to distribute detection requests is to parititon based on the content  hashing of data blocks
and the global index can also be partitioned accordingly in a  disjointed manner.
Initial data follows the VM distribution among machines and the detected results 
represented as  duplicate summary need to be collected following VM distribution. 
Therefore, there are two all-to-all data remapping operations involved.
One is to map detection requests from VM-based requests to the fignerprint based distribution.  
One is to map dupliate summary from fingerprint-based requests to the VM based distribution.  
The redistribued data needs to be accumulated on the disk to minimize the use of memory.
To minimize the cost of disk seek, outgoing or incoming data exchange messages need to be buffered to bundle small messages.
Given there are $p\times q$ partitions where $p$ is the number of machines and $q$ is the number of fingerprint-based partitions
at each machine, the number of buffers is big and space per each buffer  is small given the memory constraint.
Then there will be a large number of storage  IO operations involved, suffering the huge seek overhead.
We have designed an effient data exchange and disk data buffering  scheme to address this.

@@

{\bf Snapshot deleteion.} Each virtual machine will keep at most $x$ snapshots of VM images and thus
expired snapshots will be deleted unless its owern explicitly asks for its retention. 
A block or the entire segment can be deleted if its reference counter is zero.
%As we discussed in Section~\ref{sect:data}, the reference counter is kept in the global index, partitioned among machines. 
We perioidcally read all receipts of active 
snapsots and compute the counter of all blocks while still following the fingerprint-based partitioning.
That is based on the idea of mark-and-sweep~\cite{MarkSweep} while 
we deal such snapshot deletion in three partition-based steps similar to the VM snapshot backup.
Step 1 is to read  the meta data of snapshots such as segment and block
recipes and  accumate  reference count requests in different machines  in the fingerprint based distribution.
The second step  is to count references within each partition and detect those records with zero 
reference. The third step is that every VM gathers the confirmed deletion requests for its cotent and sends them
to the backup storage.
The archival data repository may log these deletion requests,  and will periodically perform a compaction operation when 
its deletion log is too big. 
Due to the paper constraint, we will not discuss the deletion operations in details.

%The above method requires the counter information to be consistently maintained in the global index.
%Finally the deleteion summary. it will adjust the corresponding reference counts. At the end of bucket worker's operation, it outputs the unused blocks information to each VM data repository's deletion log.

### 2013-03-08 hotstorage13/background.tex
5 papges

In a virtualized cloud environment such as ones provided by Amazon EC2\cite{AmazonEC2} and Alibaba Aliyun\cite{Aliyun}, each instance of a guest operating system runs on a virtual machine, accessing virtual hard disks represented as virtual disk image files in the host operating system.
For VM snapshot backup, file-level semantics are normally not provided.
Snapshot operations are taken place at the virtual device driver level, which means no fine-grained file system metadata can be used to determine the changed data. Only raw access information at disk block level are provided. 
Each physical hosts many VMs and a cloud cluster has  petabytes of data that need to be backed up at daily basis, 
if not more frequently. But at the same time snapshot tasks must not affect the normal cloud service, 
which means that ideally only a very small slice of CPU and memory can be used for the backup purpose.

The previous work for storage backup has extensively used  data deduplication techniques can eliminate redundancy globally among different files from different users.
Backup systems have been developed to use content hash (finger prints) to identify duplicate
content~\cite{venti02,Rhea2008}.
%,NGmiddleware2011}.
Today's commercial data backup systems (e.g. from EMC and NetApp)
%\cite{emc_avamar}\cite{datadomain_whitepaper}
use a variable-size chunking algorithm to detect duplicates in file data~\cite{similar94,hydrastor09}.

Several techniques have been proposed to speedup searching of duplicate
content. For example,
Zhu et al.~\cite{bottleneck08} tackle it
by using an in-memory Bloom filter and prefetch groups of chunk IDs that are likely to be
accessed together with high probability. It takes significant memory resource for filtering and caching.
NG et al.~\cite{ NGmiddleware2011}  use
a related filtering technique for integrating deduplication in Linux  file system and the memory
consumed is up to 2GB for a single machine. That is still too big in our context discussed below.
Lillibridge et al.~\cite{sparseindex09} break list of chunks
into large segments, the chunk IDs in each incoming segment are sampled and the segment is
deduplicated by comparing with the chunk IDs of only a few carefully selected backed up segments.
These are segments that share many chunk IDs with the incoming segment with high probability.
Deepavali et al.~\cite{extreme_binning09}  use fingerprint-based file similarity  and group similar files into the same physical location (bins) to deduplicate against each other.
That leads  to a smaller amount of memory usage for storing meta data in fingerprint
lookup with a tradeoff of the reduced recall ratio.

Additional inline deduplications are studied in ~\cite{idedup}. All of the above appraoches have focused on
such inline duplicate detection, namely given a content chunk, how to detect if it is duplicate instantantly
to trigger the data backup of this chunk or not before the detection request for the next chunk can be launched. 
In our work, this constraint is relaxed and 
there is a waiting time for many duplicate detection requests. Our goal is to minimizing  computing  resource without
affeting the overall time of backup.

>>>>

%In a virtualized cloud environment such as ones provided by Amazon EC2\cite{AmazonEC2} and Alibaba Aliyun\cite{Aliyun}, 
At a cloud cluster node, each instance of a guest operating system runs on a virtual machine, accessing virtual hard disks 
represented as virtual disk image files in the host operating system.
For VM snapshot backup, file-level semantics are normally not provided.
Snapshot operations are taken place at the virtual device driver level, which means no fine-grained file system metadata can be used to determine the changed data. Only raw access information at disk block level are provided. 
Each physical hosts many VMs and a cloud cluster has  petabytes of data that need to be backed up at daily basis, 
if not more frequently. But at the same time snapshot tasks must not affect the normal cloud service, 
which means that ideally only a very small slice of CPU and memory can be used for the backup purpose.

The previous work for storage backup has extensively used  data deduplication techniques can eliminate redundancy globally among different files from different users.
Backup systems have been developed to use content hash (finger prints) to identify duplicate
content~\cite{venti02,Rhea2008}.
%,NGmiddleware2011}.
%Today's commercial data backup systems (e.g. from EMC and NetApp)
%\cite{emc_avamar}\cite{datadomain_whitepaper}
%use a variable-size chunking algorithm to detect duplicates in file data~\cite{similar94,hydrastor09}.
Several techniques have been proposed to speedup searching of duplicate
content. For example,
Zhu et al.~\cite{bottleneck08} tackle it
by using an in-memory Bloom filter and prefetch groups of chunk IDs that are likely to be
accessed together with high probability. It takes significant memory resource for filtering and caching.
%NG et al.~\cite{ NGmiddleware2011}  use
%a related filtering technique for integrating deduplication in Linux  file system and the memory
%consumed is up to 2GB for a single machine. That is still too big in our context discussed below.
%Lillibridge et al.~\cite{sparseindex09} break list of chunks
%into large segments, the chunk IDs in each incoming segment are sampled and the segment is
%deduplicated by comparing with the chunk IDs of only a few carefully selected backed up segments.
%These are segments that share many chunk IDs with the incoming segment with high probability.
The approximation techniques are studied in Deepavali et al.~\cite{extreme_binning09} and
Zhang et al.~\cite{WeiZhangIEEE}  to reduce memory requirement with a tradeoff of the reduced deduplication ratio.
%use fingerprint-based file similarity  and group similar files into the same physical location (bins) to deduplicate against each other.
%That leads  to a smaller amount of memory usage for storing meta data in fingerprint
%lookup  
In comparison, this paper focuses on  full deduplication without approximation.

Additional inline deduplications are studied in ~\cite{sparseindex09,idedup}. All of the above appraoches have focused on
such inline duplicate detection, namely given a content chunk block, how to detect if it is duplicate instantantly
to trigger the data backup of this block or not before the detection request for the next blocks can be launched. 
In our work, this constraint is relaxed and 
there is a waiting time for many duplicate detection requests. Our goal is to minimizing  computing  resource without
affeting the overall time of backup.

### 2013-03-08 hotstorage13/intro.tex
5 papges

%Virtualization is the engine behind many popular cloud computing platforms.
%Amazon, Alibaba,  and many others have provided public VM clouds that host
%tens of thousands of  virtual machines(VM) created for various active users.
Periodic  archiving of virtual machine (VM) snaps is important for a public cloud service provider
for long-term data  retention and fault recovery in case of machine failures.  
While VM backup  may  be mainly requested by users  in some providers,  
in the case of Alibaba which provides the largest public cloud service in China,
daily backup of VM images  is conducted automatically for all  active users 
to improve data reliability and availability.

The cost of frequent backup of VM snapshots is  high because of the huge storage demand.
This issue has been addressed by   storage data deduplication ~\cite{venti02,bottleneck08} that
identifies redundant content duplicates among snapshots.  On the other hand, performing
deduplication adds significant  memory cost for comparison of content fingerprints. 
Since each physical machine in a cluster  hosts many VMs, memory contention happens frequently. 
Cloud providers often wish that the backup service only consumes  the minimal resource usage without 
impacting any existing cloud service.  Another challenge for backup with deduplication is that deletion 
of old snapshot competes computing resource with noticable cost. That is because data dependence created 
by duplicate relationship among snapshots  adds processing complexity, especially when  
VMs can migrate around in the cloud.

@@

%In this paper we intorduce the deduplication scheme of snapshot storage in Aliyun's VM cloud. We exploit the 
The rest of the paper is arranged as follows. 
Section ~\ref{sect:background}
discusses on the background and related work.
Section ~\ref{sect:arch} presents our snapshot backup solution with request and index partitioning.
Section ~\ref{sect:analysis} summarizes  our analysis of the system performance.
Section ~\ref{sect:exper} presents  our evaluation results on the effectiveness.
Section ~\ref{sect:final} concludes this paper.
% talks about Aliyun's VM snapshot system and deduplication process. In section 4 we present
%the snapshot deduplication results of our real user VMs.

>>>>

%Virtualization is the engine behind many popular cloud computing platforms.
%Amazon, Alibaba,  and many others have provided public VM clouds that host
%tens of thousands of  virtual machines(VM) created for various active users.
Periodic  archiving of virtual machine (VM) snaps is important 
for long-term data  retention and fault recovery in case of machine failures.  
For example, daily backup of VM images  is conducted automatically 
at Alibaba which provides the largest public cloud service in China.
The cost of frequent backup of VM snapshots is  high because of the huge storage demand.
This issue has been addressed by   storage data deduplication ~\cite{venti02,bottleneck08} that
identifies redundant content duplicates among snapshots.  On the other hand, performing
deduplication adds significant  memory cost for comparison of content fingerprints. 
Since each physical machine in a cluster  hosts many VMs, memory contention happens frequently. 
Cloud providers often wish that the backup service only consumes  the minimal resource usage without 
impacting any existing cloud service.  Another challenge for backup with deduplication is that deletion 
of old snapshot competes computing resource with noticable cost. That is because data dependence created 
by duplicate relationship among snapshots  adds processing complexity, especially when  
VMs can migrate around in the cloud.

@@

%In this paper we intorduce the deduplication scheme of snapshot storage in Aliyun's VM cloud. We exploit the 
\comments{
The rest of the paper is arranged as follows. 
Section ~\ref{sect:background}
discusses on the background and related work.
Section ~\ref{sect:arch} presents our snapshot backup solution with request and index partitioning.
Section ~\ref{sect:analysis} summarizes  our analysis of the system performance.
Section ~\ref{sect:exper} presents  our evaluation results on the effectiveness.
Section ~\ref{sect:final} concludes this paper.
}
% talks about Aliyun's VM snapshot system and deduplication process. In section 4 we present
%the snapshot deduplication results of our real user VMs.
